---
title: "Modeling Task 1 Report"
author: "Yash Gupta"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: false
    df_print: paged
    self_contained: true
    code_folding: hide    
  pdf_document:
    keep_tex: true
    toc: false
    fig_caption: true
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = ifelse(knitr::is_html_output(), TRUE, FALSE), 
  warning = FALSE, 
  message = FALSE,
  fig.align = "center",
  fig.width = 6,
  fig.height = 4
)
library(shiny)
library(dplyr)
library(lmtest)
library(ggplot2)
library(stargazer)
```

# **Introduction**
In the Modeling Task 1, I have been provided with U.S. Economical Quarterly Data (1950Q1 - 2000Q4) comprising of key macro-economical variables such as gdp, consumption, disposable personal income, unemployment rate etc. The primary goal was to verify if changes in people’s income (DPI) and unemployment rates can explain changes in how much they spend (consumption). I have done the following tasks in the order as mentioned below :  

1. **Verification of the specified linear model and evaluation of model accuracy** 
2. **Outlier detection using Tukey's test**
3. **Impact analysis of doubling the dataset**
4. **Autocorrelation check using Durbin-Watson test**

# **Task 1: Verification of Specified Linear Model**

## **Model Specification**

The regression model is specified as:  
\[
\Delta \text{Consumption}_t = \alpha + \beta_1 \Delta \text{DPI}_t + \beta_2 \Delta \text{Unemp}_t + \epsilon_t
\]

## **Data Preparation**

First differences were calculated for:  
- Consumption ($\Delta \text{Consumption}_t = \text{Consumption}_t - \text{Consumption}_{t-1}$)  
- Disposable income ($\Delta \text{DPI}_t$)  
- Unemployment rate ($\Delta \text{Unemp}_t$)  

```{r data-prep}
df <- read.csv('data/ModelingTask1.csv', na.strings = 'NA')

df_diff <- df |>
  mutate(
    D_consumption = c(NA, diff(consumption)),
    D_dpi = c(NA, diff(dpi)),
    D_unemp = c(NA, diff(unemp))
  ) |>
  filter(!is.na(D_consumption))
```

Note : Only 1 row was removed due to differencing operation.

## **Results**

```{r task1}
model <- lm(D_consumption ~ D_dpi + D_unemp, data = df_diff)
summary_model <- summary(model)

knitr::kable(coef(summary_model), digits = 4,
             caption = "Linear Model Output (Task 1)")
```

--- 

## **Model Accuracy** 

Since model was build on the whole data and any testing data was not provided so I checked the accuracy metrics for regression on the training data itself. 

```{r accuracy, include=TRUE, echo = TRUE}
# Calculating R-squared, RMSE, MAE
rsq <- summary_model$r.squared
residuals <- residuals(model) 
rmse <- sqrt(mean(residuals^2))
mae <- mean(abs(residuals))
```

```{r ,echo = FALSE, include=TRUE}
cat("Accuracy Metrics:\n",
    "R-squared:", round(rsq, 3), "\n",
    "RMSE:", round(rmse, 2), "\n",
    "MAE:", round(mae, 2))


```


